---
title: "Project 2"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

## Question 1: 
```{r}
bank_full <- read.csv(file = '/Users/victortu/Desktop/ BUS 461/BUS-461A/Csv files/bank-full.csv', sep = ";" )
summary(bank_full)
```

After looking at the output for summary(bank_full) it looks like the summary function gives us very useful and quick insight into our dataset.    
Key takeaways:    
-For columns that are numerical, we get statistics on them such as the min, 1st quartile, median, mean, 3rd quartile, and max. These are extremely useful statistics for when we need to make quick analyses of the columns in the dataset.    
-For columns that are categorical, we get the length, class, and mode of the data in the column.     

## Question 2:
```{r}
summary(bank_full$campaign)
```

After looking at the summary of the "campaign" column of bank_full, I noticed that the max amount of campaigns in the dataset is 63.000, which seems pretty large. This could be an outlier because the mean and median of the entire "campaign" column is 2.764 and 2.000, respectively. The mean and median values look accurate when compared to the min, 1st quartile, and 3rd quartile values. We have a right or positive skew of values in the "campaign" column.  

## Question 3: 
```{r}
boxplot(bank_full$campaign, main = "Campaign Boxplot")
```

The box represents min, quartile 1, quartile 2, the median, and quartile 3. From the boxplot, there are quite a few outliers in the "campaign" column (about 3).

## Question 4: 
```{r}
per95 <- quantile(bank_full$campaign, + c(0.1,0.2,0.3,0.4,0.5,0.6,0.75,0.85,0.95))
boxplot(per95, main = "Campaign 95th percentile Boxplot")
```

The above boxplot represents the 95th percentile values of the "campaign" column. This box represents the same min, but it now represents a different quartile 1, 2, median, and quartile 3 (becuase we only have the 95th percentile of "campaign" values). Now there is only one outlier in the boxplot (8).The solid horizontal line in the boxplot represents the median of the 95th percentile of "campaign" values. 

## Question 5: 
Part 1:    
```{r}
hist(per95, col = "lightyellow", main = "Hist of per95")
```

The histogram above is what I would expect. The highest frequency of campaigns is 1 to 2 and the lowest frequency of campaigns is 7 to 8.    

Part 2: 
```{r}
hist(per95, col = "lightyellow",  main = "Hist of per95 with rug")
rug(per95, col = "darkred", ticksize = ".05", lwd = "0.7")
```

The histogram above is with the rug function applied. The rug function allows us to create a set of tick marks along the base of the plot (these tick marks are in dark red). These tick marks allow us to know exact values. 

## Question 6: 
```{r}
hist(per95, col = "lightyellow", breaks = 25, main = "Hist of per95 with breaks = 25 and abline") 
abline(v = median(bank_full$campaign))
```

## Question 7:
```{r}
graphCountDeposits <- table(bank_full$y)
barplot(graphCountDeposits, col = "lightyellow", main = "Comparing customers with/without term deposits", ylim = c(0,40000))
```

## Question 8:
```{r}
dfmanagement <- subset(bank_full, job == "management" & balance >= "0")
dfstudent <- subset(bank_full, job == "student" & balance >= "0")

boxplot(dfmanagement$balance, main = "Management Bal")
boxplot(dfstudent$balance, main = "Student Bal")
```

From the boxplots for managmement balance and student balance, we can see that the balance of a management, on average, is larger than students. Both box plots have outliers and follow the same trends.

## Question 9 
```{r}
plot(bank_full$campaign, bank_full$duration)
```

## Question 10
```{r}
plot(bank_full$campaign, bank_full$duration, xlab = "Phone call length (min)", ylab = "Number of times a customer is called", main = "Duration vs Campaign")
```

Observations:   
-Calls with a shorter duration have an increased number of times a customer is contacted, and vice versa     
-The outliers are at the bottom right of the scatterplot and top left     
-In Question 2, I mentioned that the "campaign" column values are rigth skewed, this seems to be the same case for this graph as well  